---
title: "Wuhan: Seeding vs transmission"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
```


```{r, message = FALSE}
library(WuhanSeedingVsTransmission)
library(fst)
library(tidyr)
library(tidyr)
library(purrr)
library(knitr)
```

# Scenarios and parameter sampling

* Stratify the analysis by seeding event size and duration:
  * Event size: 20, 40, 60, 80, 100
  * Event duration: 7 days, 14 days, 28 days
* Sampled parameters:
  * Serial interval: 8.4 (SD: 3.8) from Lispsitch et al. (2003)
  * K: 0.16 from Lloyd-Smith (2005)
* Fitted parameters
  * R0: Uniform(0, 5) prior
* Case bounds:
  * Lower bound: 780
  * Upper bound 880

# Scenario analysis

1. Branching process from `bpmodels`.
   * Assumes a negative binomial distribution on new cases
   * Assumes that the serial interval is gaussian distributed. 
2. Run an outbreak for each day of the assumed seeding event.
3. Summarise outbreaks across a single seeding event.



# Restrict scenarios and summarise output

* Restrict potential scenarios based on an upper and lower bound of total cases.

* Summarise each outbreak across samples.

* Report number of accepted samples per scenario.

* Condition on reported cases on the 3rd and the 20th of January


```{r read-in-results}
scenario_sims <- fst::read_fst("../inst/results/grid.fst")
```

```{r}
## Filter by known cases on the 3rd of January
jan3_conditioned <- WuhanSeedingVsTransmission::condition_and_report_on_cases(
                                                  scenario_sims, 
                                                  condition_date = "2020-01-03",
                                                  samples = max(scenario_sims$sample),
                                                  lower_bound = 40, 
                                                  upper_bound = 400)
  

## Filter by known cases on the 20th of January
jan20_conditioned <- WuhanSeedingVsTransmission::condition_and_report_on_cases(
                                                  jan3_conditioned$conditioned_sims, 
                                                  condition_date = "2020-01-20",
                                                  samples =  max(scenario_sims$sample),
                                                  lower_bound = 200, 
                                                  upper_bound = 2000)



## Filter by known cases on the 25th of January
jan25_conditioned <- WuhanSeedingVsTransmission::condition_and_report_on_cases(
                                                  jan20_conditioned$conditioned_sims, 
                                                  condition_date = "2020-01-25",
                                                  samples =  max(scenario_sims$sample),
                                                  lower_bound = 1400, 
                                                  upper_bound = 4000)
```

# Output

Compare R0s across scenarios (reporting median, maximum and minimum):

* Grid of event size vs event duration with R0 estimates.


```{r plot-probs}
WuhanSeedingVsTransmission::plot_scenario_tile(jan25_conditioned)
```

```{r summarise_r0s}
## Make current r0 summary based on current time
summarised_r0s <- jan25_conditioned$conditioned_sims %>% 
  dplyr::filter(upper_R0 == 4) %>% 
  dplyr::group_split(serial_mean) %>% 
  purrr::map(  ~ WuhanSeedingVsTransmission::summarise_end_r0(.) %>% 
  WuhanSeedingVsTransmission::make_duration_size_table() %>% 
  knitr::kable()) 
```

```{r low_serial_r0}
summarised_r0s[[1]]
```

```{r high_serial_r0}
summarised_r0s[[2]]
```